---
title: "introduction"
author: "John Palmer"
date: "23/01/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Hey! Figured I'd create this because there was genuinely no introduction to base R usage in our course. I figured this might be easier than typing a big long text explanation. As an alternative, you can also read through the lengthier R markdown tutorial at [this link](https://rmarkdown.rstudio.com/lesson-1.html) if you want.

### R Markdown

I'll briefly introduce R Markdown (Rmd; the format you're viewing this in). 
If you already feel comfortable, by all means skip to the `Assignment` section. 
So, what's the point of it? Rmd allows you to generate documents that seamlessly combine text with snippets of code. 
To view the fully formatted document, click "knit" at the top of RStudio (makes for some nice reading).
Also, HTML format should always work, while PDF might require LaTeX to be installed. 

Without any special formatting, the default format is standard text like what I'm writing in this sentence. 
Now this "standard text" is actually markdown language, meaning there are various symbols I can use to control text formatting. 
You probably already picked up on my use of titles (more `#` symbols means smaller) and paragraphs (every empty line starts a new one). 
Some of this stuff isn't that important, like **bolding**, *italics*, etc., so I'll just leave [this link](https://www.markdownguide.org/basic-syntax/) in case you want to learn more.


To complete the assignment, the actual code formatting is what you really want. 
Definitely the most important is triple backquotes (not single or double quotes, it's the symbol on the same key as the tilde (~) symbol). 
The first three quotes mark the start, and the second three mark the end (meaning you can add/remove cells if needed).

```
like this
```
But what I wrote is only recognized as plain text, because Rmd needs to know what language you're using.
This is why you see `{r}` next to the starting backquotes (so that it recognizes it as R code). 

```{r highlevelmath}
num1 <- 17
num2 <- 4
num1 + num2
```

Specifically, the first "r" tells it to expect R code, and the word "highlevelmath" after is just a label.
In case you haven't seen it, the green 

You can also write code on the same line as text using single backquotes. 
Same as before, if you specify the "r" flag, you can also run R code like this: `r 2*3`, or even using variables: `r myvariable = 27` `r myvariable + 3`.
Without the "r" flag, it is still quite handy for emphasizing a variable or name, like this: `hey im important`.


### Assignment

First, searching up function help pages is essential. All you need to do is throw a `?` in front of the function you want to learn more about. 

```{r}
?qqnorm()
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
