---
title: "R Intro"
author: "John Palmer"
date: "23/01/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hey! Figured I'd create this because there was genuinely no introduction to base R usage in our course. I figured this might be easier than typing a big long text explanation. Here, I'll try and show you the tools I used to answer the questions.

### Question 1
First, you'll need to get comfortable searching up function help pages. All you need to do is throw a `?` in front of the function you want to learn more about. 

```{r}
?rnorm
```
There's some real good documentation online too if the manuals aren't clear. So, what does `rnorm(100)` do? It randomly draws 100 numbers from a normal distribution (got that from the help page). 

### Question 2

Here are some common techniques regarding vectors and loops:
```{r}
vect <- 1:20        # Quick way to create a vector of all numbers from 1 to 20
vect

# Using a loop
vect <- c()                      # Initialize an empty vector, c is the letter used for vector, no clue why
for(i in 1:20){        
  vect[i] <- i                   # Load the vector
}
vect

# Adding up all the numbers in vect
sum <- 0
for (i in 1:length(vect)){        # Using length() is a really common technique for traversing an existing vector
  sum <- sum + vect[i]            
}
sum

sum(vect)           # Or using the sum function

# Getting the maximum value from "vect"
max <- -Inf                       # I choose the lowest number possible
for (i in 1:length(vect)){        
  if (vect[i] > max){
    max <- vect[i]
  }
}
max

max(vect)         # Or using the max function

```


### Question 3 
R is also pretty smart about performing math on vectors or other large structures. 
This is a REALLY important (and also confusing) topic in R. 
```{r}
vect          # Vect is still saved in memory from the previous cell!

vect * 2 

1/vect

log(vect)

log(vect) + 5
```

Here's a relevant example:
```{r}
myVector <- 1 / (1:500)         # This vector is massive, so I display only the first 20 results using head()
head(myVector, 20)

sum(myVector)                   # the sum from i=1 to 500 of 1/i (hint)

# Load the vector 
N <- c(500, 1000, 2000, 4000, 8000)

# A loop might help here!
results <- c()
for (i in 1:length(N)){
  results[i] # <- SOMETHING      ## Calculate something using N[i] to fill the "results" vector 
}


```



### Question 4 : Matrix 

```{r} 
mat <- matrix(nrow=4, ncol=8)     # Initialize an empty matrix (heads up, not the correct dimensions)
mat

# Load the 2nd column      ## Notice that R is ALWAYS [row,column] (for other things too)
mat[,2] <- c(1,2,3,4)      ## Leaving a row or column empty signifies all
mat

# Load the 3rd row
mat[3,] <- 8:1           
mat             
```
Here I'll just throw in some examples of `rep()`.
The two key arguments you need are `times` and `each`. 
`times` tells you how many times the whole vector gets repeated. 
`each` tells you how many times each individual element within the vector gets repeated. 
```{r rep}
rep(c(1,2), times=4)      # Times

rep(c(1,0), each=4)       # Each

rep(c(2,4), each=4, times=2)    # Both
```

### Question 5
Data frames are probably the most important structure in all of R. They're insanely useful.
They are kind of like a matrix where each column can hold a different type of data (matrix is all numeric I believe).
The same "row,column" format applies to data frames too.
```{r}
data("Formaldehyde")      # This is used to load the formaldehyde data frame
class(Formaldehyde)       # Just showing its a DF
Formaldehyde
```
Referencing stuff in a data frame:
```{r}
# Many different ways
Formaldehyde$optden         # Using the "$" operator
Formaldehyde[,2]            # Entire 2nd column           
Formaldehyde[,'optden']     # Entire "optden" colum     # Notice that all 3 ways are exactly the same

Formaldehyde[4,]            # Entire 4th row

```

### Question 6 & 7 
The use of boolean vectors in R is extremely powerful. By boolean vector, I mean a vector of True/False values.
```{r}
1:50 %% 2 == 0           # An example of a boolean vector 
```


```{r}
library(MASS)               # This data frame is only found in the MASS package, so I need to import it 
data("Pima.tr")             # Data() is used to retrieve a data source
head(Pima.tr)

Pima.tr$age > 40  # This boolean has the same number of entries as the number of rows in Pima.tr
```

When I plug this boolean vector into the "row" field of the data frame, it selects all the rows corresponding to "True" in the vector. Crazy right?
```{r}
myBool <- Pima.tr$age > 40
Pima.tr[myBool,]         # Remember [row,column]
```
Anyways, I hope that helped! I figured I'd do this because I think learning the basics of R without any reference material would suck major butt (I got major help from a post-doc researcher in my lab). Cheers!