---
title: "R Markdown Introduction"
author: "John Palmer"
date: "23/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Greetings
Hey there! Here I'll just cover the basics of R Markdown quickly. As an alternative, you can also read through the lengthier R markdown tutorial at [this link](https://rmarkdown.rstudio.com/lesson-1.html) if you want.

### R Markdown

I'll briefly introduce R Markdown (Rmd; the format you're viewing this in). 
So, what's the point of it? Rmd allows you to generate documents that seamlessly combine text with snippets of code. 
To view the fully formatted document, click "knit" at the top of RStudio (makes for some nice reading).
Also, HTML format should always work, while PDF might require an additional package (LaTeX) to be installed. 

Without any special formatting, the default format is standard text like what I'm writing in this sentence. 
Now this "standard text" is actually markdown language, meaning there are various symbols I can use to control text formatting. 
You probably already picked up on my use of titles (more `#` symbols means smaller) and paragraphs (every empty line starts a new one). 
Some of this stuff isn't that important, like **bolding**, *italics*, etc., so I'll just leave [this link](https://www.markdownguide.org/basic-syntax/) in case you want to learn more.


To complete the assignment, the actual code formatting is what you really want. 
Definitely the most important is triple backquotes (not single or double quotes, it's the symbol on the same key as the tilde (~) symbol). 
The first three quotes mark the start, and the second three mark the end (meaning you can add/remove these cells whenever you want).

```
like this
```
But what I wrote above is only recognized as plain text, because Rmd needs to know what language you're using.
This is why you see `{r}` next to the starting backquotes (so that it recognizes it as R code). 
When you include this bit of formatting, it also includes a built-in "run" button (little green play button) next to the cell. 
Really handy to check your result for a single cell while you're editing in RStudio. 

```{r highlevelmath}
num1 <- 17
num2 <- 4
num1 + num2
```

In the above cell, the first "r" tells it to expect R code, and the word "highlevelmath" after is just a label.

You can also write "inline" code using single backquotes. 
Same as before, if you specify the "r" flag, you can run R code like this: `r 2*3`, or even using variables `r myvariable = 27` that Rmd remembers: `r myvariable + 3`.
Without the "r" flag, it will not run as code, but it's still quite handy for emphasizing a variable or name, like this: `hey im important`.

You can also check out this cell. I'm just using it to show how one cell can produce multiple outputs (in various formats).
```{r}
summary(cars)
plot(cars)
```
